A Geometry for Network Visualization in \MakeLowercase{\texttt{ggplot2}}

### 2015-04-01 CC/writeup.Rnw
conclusions/future work

  \subsection{Other Arguments}
  Similar to the other geometries and functions in \texttt{ggplot2}, our geometry takes several arguments outside of the aesthetic mapping parameters.  One such argument is the \texttt{vlabel} argument. The \texttt{vlabel} aesthetic is a logical value that, if set to \texttt{TRUE} prints the name or number associated with each vertex in the \texttt{label} column in order to identify it on the plot.  In the final version of this geometry, we will add capabilities to adjust the size and positioning of the labels. For the moment, the color is set to the vertex color, the size is set to twice the vertex size, and all other possible parameters are set to their defaults in \texttt{geom\_text()}. The \texttt{vlabel} argument is most useful for smaller network objects where all vertex names can be printed on their corresponding vertices and still be read clearly.  Next, the \texttt{directed} argument is a logical value that identifies whether or not the network is directed.  When this value is true, arrows are created on the end of the line segment that corresponds to the \texttt{to\_id} value for each edge.  The arrows match the edges in coloring, linetype, and size, and the default length of the arrow is set using the \texttt{unit} function in the \texttt{grid} package to 0.015 of the normalised parent coordinates of the plot.  
  Finally, the \texttt{layout} argument takes a character value corresponding to the possible layouts in the \texttt{sna} package that are created by the \texttt{gplot.layout.*()} family of functions.  The default layout is the Kamada-Kawai layout.  This is a force-directed layout for undirected networks \citep{kamadakawai}. There are, however, many other layouts possible.    
\begin{table}
\centering
\begin{tabular}{l|p{10cm}}
Layout Name & Description \\
\hline \hline
\texttt{"adj"} & a version of \texttt{"mds"} which scales the raw adjacency matrix \\
\texttt{"circle"}* & places vertices uniformly in a circle \\
\texttt{"circrand"} & places vertices randomly in a circle \\ 
\texttt{"eigen"} & places vertices based on the eigenstructure of the adjacency matrix \\
\texttt{"fruchermanreingold"} & uses a variant of Fruchterman and Reingold's force-directed placement algorithm \\
\texttt{"geodist"} & a version of \texttt{"mds"} which scales the matrix of geodesic distances \\
\texttt{"hall"}* &  places vertices based on the last two eigenvectors of the Laplacian of the input matrix \\
\texttt{"kamadakawai"} & generates a vertex layout using a version of the Kamada-Kawai force-directed placement algorithm \\
\texttt{"mds"}  & places vertices based on a metric multidimensional scaling of a specified distance matrix \\
\texttt{"random"} & places vertices randomly \\
\texttt{"segeo"} & a version of \texttt{"mds"} which scales the squared euclidean distances between row-wise geodesic distances \\
\texttt{"seham"} & a version of \texttt{"mds"} which scales the Hamming distance between rows/columns of the adjacency matrix \\
\texttt{"target"} & produces a ``target diagram" or ``bullseye" layout using a force-directed placement algorithm.
\end{tabular}
\caption{All possible layouts to pass to \texttt{geom\_net}.  All descriptions are from \citet{sna}.  Note that the layouts marked with an asterisk (*) are the only layouts that do not take layout parameters.}
\label{table:layouts}
\end{table}

@@

\section{Future Work}
As we discussed throughout the paper, there is still some work to be done before we submit it for publication in \emph{The R Journal}.  We also hope to make it an official part of \texttt{ggplot2}.  We believe that this is an important contribution to \texttt{ggplot2} and to the R user community because it fills in gaps in the network plotting capabilities of R. \sct{XXX That's all I got. Help?}

>>>>

  \subsection{Other Arguments}
  Similar to the other geometries and functions in \texttt{ggplot2}, our geometry takes several arguments outside of the aesthetic mapping parameters.  One such argument is the \texttt{vlabel} argument. The \texttt{vlabel} aesthetic is a logical value that, if set to \texttt{TRUE} prints the name or number associated with each vertex in the \texttt{label} column in order to identify it on the plot.  In the final version of this geometry, we will add capabilities to adjust the size and positioning of the labels. For the moment, the color is set to the vertex color, the size is set to twice the vertex size, and all other possible parameters are set to their defaults in \texttt{geom\_text()}. The \texttt{vlabel} argument is most useful for smaller network objects where all vertex names can be printed on their corresponding vertices and still be read clearly.  Next, the \texttt{directed} argument is a logical value that identifies whether or not the network is directed.  When this value is true, arrows are created on the end of the line segment that corresponds to the \texttt{to\_id} value for each edge.  The arrows match the edges in coloring, linetype, and size, and the default length of the arrow is set using the \texttt{unit} function in the \texttt{grid} package to 0.015 of the normalised parent coordinates of the plot.
  Finally, the \texttt{layout} argument takes a character value corresponding to the possible layouts in the \texttt{sna} package that are created by the \texttt{gplot.layout.*()} family of functions.  The default layout is the Kamada-Kawai layout.  This is a force-directed layout for undirected networks \citep{kamadakawai}. There are, however, many other layouts possible.
\begin{table}
\centering
\begin{tabular}{l|p{10cm}}
Layout Name & Description \\
\hline \hline
\texttt{"adj"} & a version of \texttt{"mds"} which scales the raw adjacency matrix \\
\texttt{"circle"}* & places vertices uniformly in a circle \\
\texttt{"circrand"} & places vertices randomly in a circle \\
\texttt{"eigen"} & places vertices based on the eigenstructure of the adjacency matrix \\
\texttt{"fruchtermanreingold"} & uses a variant of Fruchterman and Reingold's force-directed placement algorithm \\
\texttt{"geodist"} & a version of \texttt{"mds"} which scales the matrix of geodesic distances \\
\texttt{"hall"}* &  places vertices based on the last two eigenvectors of the Laplacian of the input matrix \\
\texttt{"kamadakawai"} & generates a vertex layout using a version of the Kamada-Kawai force-directed placement algorithm \\
\texttt{"mds"}  & places vertices based on a metric multidimensional scaling of a specified distance matrix \\
\texttt{"random"} & places vertices randomly \\
\texttt{"segeo"} & a version of \texttt{"mds"} which scales the squared euclidean distances between row-wise geodesic distances \\
\texttt{"seham"} & a version of \texttt{"mds"} which scales the Hamming distance between rows/columns of the adjacency matrix \\
\texttt{"target"} & produces a ``target diagram" or ``bullseye" layout using a force-directed placement algorithm.
\end{tabular}
\caption{All possible layouts to pass to \texttt{geom\_net}.  All descriptions are from \citet{sna}.  Note that the layouts marked with an asterisk (*) are the only layouts that do not take layout parameters.}
\label{table:layouts}
\end{table}

@@

\section{Future Work}
\hh{XXX  just be a bit more specific here: we (a) want to make the code a part of the ggplot2 world - either as a part of ggplot2 itself or within GGally. (b) we want to publish this paper in the R Journal.
Things that we still need to do are discussed in the paper, they are: .... }
As we discussed throughout the paper, there is still some work to be done  before we submit it for publication in \emph{The R Journal}.  We also hope to make it an official part of \texttt{ggplot2}.  We believe that this is an important contribution to \texttt{ggplot2} and to the R user community because it fills in gaps in the network plotting capabilities of R. \sct{XXX That's all I got. Help?}
